### 2020-01-02 Literature Review/Literature Review - Revised/sections/MARL.tex

\documentclass[../sample.tex]{subfiles}

\begin{document}

Reinforcement learning extends the Markov Decision Process problem by considering the case where the
reward model is not initially known to the agent. In a similar manner, Multi Agent Reinforcement
Learning (MARL) extends the Markov Game setting to one where the payoff structure is not a priori
knowledge.

The task of MARL is to determine an optimal joint policy for all agents across the game. This joint
policy may be the concatenation of all the individual policy or it may just be options for each
agent to take. In either case, optimality is defined through the standard notions of Nash equilibria
and so, in this section, I will try to consider the broad spectrum of methods which attempt to
achieve this Nash equilibria. The largest problem in MARL is the non-stationarity of the environment
\cite{Hernandez-LealA}. In single-agent settings, it is assumed that the environment is Markovian.
However, this must be lifted in the Multi Agent setting since other agents in the environment will
be learning concurrently. As such, we must now consider that the policy for any one agent will
depend on the policy of all other agents.

This chapter begins with a selection of the foundational methods which were developed towards
solving the MARL problem. The interested reader may find a stronger review and analysis in
\cite{SchwartzMulti-agentApproach}. I then go on to consider more contemporary approaches, most
notably that of agent modelling.

\section{Learning in Two Player Matrix Games}

The most fundamental method to learning in Matrix games is the simplex algorithm. This is a popular
method of linear programming (in which constraints are linear). This will be important in
considering more current methods. A similar consideration is given to the infinitesimal gradient
ascent algorithm, in which the step size converges to zero. This method guarantees that, in the
infite horizon limit, the payoffs will converge to the Nash equilibrium payoff. Note that this does
not necessarily mean that both agents will converge to a single Nash equilibrium. This is a
particular problem in games where there are multiple Nash equilibria. However, in practice it is
difficult to choose a convergence rate of the step size and, without an appropriate choice the
strategy may oscillate as shown in the book. To address this, a modified approach is presented by
Bowling and Veloso which incorporates the notion of Win or Learn Fast (WoLF) to produce WoLF-IGA.
WoLF is a notion we will come across often in MARL and is shown by the authors to converge to always
to a NE. The concern with WoLF methods, however, is that it requires explicit knowledge of the
payoff matrix (which is not so much of a problem for model based methods) and the opponent's
strategy (which is more of a problem in real-world methods). Finally, the Policy Hill Climbing
method (PHC) is shown to converge to an optimal mixed strategy if the other agents are stationary
(i.e. are not learning). However, it is shown that, when this is not the case, the algorithm again
oscillates. The WoLF-PHC adaptation of this method is shown to converge to a NE strategy for both
players with minimal oscillation.

\section{Learning in Multiplayer Stochastic Games}

Stochastic Games (or Markov Games) form a basis for MARL settings. However, in this case the agents
must learn about the equilibrium strategies by playing the game, which means they do not have a
priori knowledge of the reward or transition functions. Schwarz considers two properties which
should be used for evaluating MARL algorithm: rationality and convergence. The latter simply states
that the method should converge to some equilibrium whereas the former suggests that the method
should learn the best response to stationary opponents. A similar set of conditions is considered by
Conitzer and Sandholm in \cite{ConitzerAWESOME:}, whose algorithm we will consider shortly. Schwarz
then presents a review of MARL methods (as of September 2014)

\TODO{Consider the case for partially observed/decentralised settings. Find the relevant material
within \cite{SchwartzMulti-agentApproach} which addresses this issue.}

\subsection*{Agent Modelling}

Returning to the problem of non-stationarity, solutions have been presented in which the agent
models the learning of other agents. A noteworthy example of this is found in
\cite{Foerster2018LearningAwareness} in which the agent performs a one-step lookahead of the other
agents' learning and optimises with respect to this expected return. They show that this leads to
stable learning and can even lead to emergent cooperation from competition. However, the method
requires that both agents have exact knowledge of the others' value functions in order to perform
the one step lookahead. Furthermore, it has only been considered for the case of a two agent
adversarial game and so the scalability of the system to multiple agents is not yet understood.
Another method presented by Mao et al. \cite{MaoModellingDDPG} uses a centralised critic to collect
the actions and observations of all agents and allows it to model the joint policy of teammates.
This is shown to generate cooperative behaviour across four agents and so is more applicable to real
world settings. However, its disadvantage over the method presented in
\cite{Foerster2018LearningAwareness} is that the critic is centralised. In real world settings, this
requires the presence of an agent (perhaps a laptop) which is able to handle the computational load
of determining a joint policy across all agents and must then communicate the Q-values of all agents
back to them. This is both a taxing both in terms of computation and time.

Hong et al \cite{Hong2018ASystems} present a similar system for modelling teammate policies by
tasking a CNN with determining the policy features of other agents and then embedding these as
features in its own DQN. This shows strong performance in settings where other agents dynamically
change their policies. The concerns with this, however, are that, as the number of agents in the
field increase, the CNN in each agent must perform another approximation. This places strong
requirements on the performance of the CNN since errors in estimation will accumulate as the number
of agents increases. Similarly, the complexity of the DQN will increase as more feature vectors are
added.

Finally, all of the above methods are not robust to evolving numbers of agents. The problem of agent
modelling is an important one to ensure stable learning and to understand the evolution of the
system. It also presents a strong challenge and is open to exploration. To put it in context the
methods described in this section are all from 2018-19, so its all very new.

The above methods are all centralised techniques, in which a controller determines the optimal joint
policy for both agents. However, in real scenarios it is often preferable that each agent learns
their own strategy, a task which must be completed without information of the other agent's
strategy. The methods presented towards this problem are: linear reward-inaction ($L_{R-I}$) which
guarantees convergence to NEs in games which contain pure NEs, linear reward-penalty ($L_{R-P}$)
which can guarantee convergence to mixed strategies given the appropriate parameters, lagging anchor
algorithm which also converges to mixed strategies, and the author's own proposal of the $L_{R-I}$
lagging anchor algorithm which can converge to both pure and mixed NEs.

\section{Stable Learning}

This section considers a dynamical systems approach towards learning agents. Its aim is to develop learning systems which prioritise stability.

Stability may be looked at from the view point of two perspectives. The first is from an
optimisation point of view. This considers the dynamics of the learning model, allowing us to better
choose our parameters and design our models so that they may converge to a stable result. The second
is from the view point of the state-action space of a learnt model. This allows us to determine,
before the MAS is deployed, which set of state-action pairs will lead to unstable behaviour. This
knowledge allows us to consider which state-action pairs should be avoided. In both cases, stability
analysis allows us to build multi agent systems which will learn and act in the way that we expect
them to.

In \cite{Letcher2019DifferentiableMechanics}, Letcher et al. model gradient descent learning of
generative-adversarial-networks (GANs) as a two-player differentiable game. A differentiable game is
one in which the loss function is twice differentiable. Using this formulation, they are able to
analyse the system from new perspectives by considering the current state-of-the-art understanding
of differentiable game theory. Whilst, at first glance, this may seem like a purely theoretical
exercise, they go on to show that the insights gained allow them to develop a new multi-objective
optimisation technique for GANs which shows stronger convergence properties, most notably of which
is that it guarantees that the method finds a stable equilibrium (and avoids saddles) between the
two players' loss functions.

Jin and Lavaei \cite{Jin2018Stability-certifiedPerspective} consider the policy of a reinforcement
learning agent as a non-linear, time varying feedback controller. Using this notion, they then
consider the bounded-input-bounded-output stability of the system. They do this by analysing the
ratio between the total output and total input energy (called the L2 gain). If the L2 gain remains
finite then the system may be considered to be stable. They then apply these considerations on
real-world applications including multi-agent flight formation and obtain stability certificates
(essentially confirming that the system will remain stable under certain conditions) for the learned
controller.

Berkenkamp et al. \cite{Berkenkamp2017SafeGuarantees} consider a similar problem from a different
definition of stability. Specifically, they look at stability from the point of view of Lyapunov
functions. A system is said to be stable if the applying the policy will result in stricly lower
evaluations of the function. In other words, a system is stable if its corresponding Lyapunov
function is decreasing towards a minimum point. The authors use this idea to define a 'region of
attraction' in which the system is stable in the sense of Lyapunov. The goal of Safe Lyapunov
Learning, a method which they develop from these insights, is to learn a policy without leaving this
region of attraction. They do this by taking measurements within the set and using this to learn
about the system dynamics, thereby increasing the safe set. With this, we now have a guarantee that
policy optimisation will not result in unsafe behaviour, even in the presence of stochasticity and
exploration.

As we can see from the discussion thus far, there are many definitions and perpectives of stability,
each lending to a new understanding of learning systems. To add another to the mix, Milchtaich in
\cite{Milchtaich2007StaticGames}, presents a notion of static stability. This means that it is based
solely on the incentives of the players and does not require a consideration of the dynamics of the
system. Milchtaich's definition of a stable system is one in which, when perturbed, it is more
beneficial for an agent to move back towards the equilibrium than it is for them to move away. This
is perhaps the most fundamental definition of a stable equilibrium and, as such, Milchtaich shows
that it is applicable to all strategic games, within certain assumptions (finite set of player and
continuous strategy space) and considers probabilistic perturbations from the original state. This
is particular applicable to Multi Robot Reinforcement Learning (or any MARL in continuous strategy
spaces) since these systems require random exploration of the strategy space and the dynamics are
not known a priori. However, the lack of dynamics means that we do not consider the evolution of
learning.

\TODO{In the above, look specifically at the assumptions made within the papers. In the multi-agent
case, the biggest assumption is that of independent learning. Perhaps consider how advances in
coupled dynamical systems may be used to consider learning agents whose actions affect one another.}

\end{document}