### 2020-01-07 Literature Review/Literature Review - Revised/sections/Proposals.tex

\documentclass[../sample.tex]{subfiles}

\begin{document}

\section{Considering Intelligence in Swarm Dynamics}

Most of the diffusion models, as mentioned in ‘Swarms’, do not take into consideration the
interactions between agents, since it adds a large degree of complexity to the model. However, it
would be important from a safety perspective to take these into account by looking at the attractive
and repulsive potentials across agents. This would allow us to provide a guarantee of inter-agent
collision avoidance whilst maintaining the progression of the swarm towards desired tasks. As
mentioned in ‘Swarms’, advances have been made towards considering the pairwise interactions between
swarm agents \cite{Bellomo2017}. Importantly, this addition presents the first step towards
considering the intelligence of the individual within the swarm.

By expanding the control perspective of stochastic swarms, predictive control methodologies can be
developed, similar to that of \cite{Borzi2015} which allow swarms to produce more complex phenomena
such as constraint satisfaction. As an example of where this might be important, consider the
problem of swarm systems in urban search and rescue, where one swarm sub-group is required to carry
a victim away from the scene, whilst another is required to search the area. This can be viewed as
from a control perspective where it is required that the latter group’s macroscopic model ensures
that agents do not enter within a safe region of the former group.

Stochasticity can then be considered from a hybrid perspective through JMLS (see Hybrid Control),
allowing the control laws to take into account agent and communication failures.
\cite{FUHRMANFrancescoRUSSO,Ma2017,Li2017} have begun to make progress in these areas, but it does
not seem like the connection of MJD and swarm dynamics has been made yet.

Finally, it may be appropriate to consider stochastic control of heterogenous swarms to allow for a
‘marriage’ between swarm dynamics and game dynamics. This presents the possibility for swarms
sub-groups to adapt to each other’s behaviour through repeatedly play and perhaps display elements
of cooperative (or non-cooperative) behaviour, all while maintaining the required properties of
stability and controllability.

\section{Stochastic Predictive Control for Decentralised Hybrid Systems}

The purpose of this method is to leverage the ideas presented in \cite{Foerster} and
\cite{Heirung2019}. Here, agents must maintain (and perhaps learn) a model of the other agents
within the system. This model provides some representation of the trajectories that the other agents
will carry out or an understanding of their future actions. Using their own system models in tandem
with these belief models, agents must then determine their trajectory. The challenge here would be
proving the stability and controllability of the system whilst maintaining minimal communication
requirements. However, it would appear to be a feasible area of exploration.

This system can then be expanded by considering the case of stochastic hybrid systems. This draws on
the ideas presented in ‘Hybrid Control’ where we argue that agents in the system may be subject to
stochastic failures. It is, therefore, required that we develop control methodologies which are
robust to these switching dynamics to ensure their safe operation.

Similarly, it would be important to consider the learning aspect of these systems. The method so far
assumes that each agent knows all possible modes of operation of other agents and their transition
probability. However, this is unlikely to be the case. As such, it would be important to consider
the ability of agents to learn these models through iterative interaction and learning. This may
improve the robustness and optimality of the system, whilst maintaining its desired properties.

On a similar topic of learning, it would be vital, from a failure perspective, to consider fault
detection in a decentralised hybrid system. The above considers learning the system model of other
agents who have undergone a faiure. However, it is equally interesting to consider learning 
\textit{that} they have failed in the first place. The results suggested in 'Swarms' may be
applicable here, in an effort to maintain the decentralisation of the system.

Finally, these methods could be extended towards settings where the resultant behaviour is not best
described by a Nash Equilibrium. The control theory literature, thus far, has shown a strong
assumption that the system will always converge to a Nash Equilibrium. However, in games of repeated
play, the resultant dynamics of the game may be best described through cyclic or recurrent behaviour
\cite{Boone2019FromTheory}. Taking this into consideration would allow distributed control to be applied in a
wider array of settings, particularly in those where agent interests are conflicting with one
another.

\section{Dynamical Study of Multi Agent Learning}

This builds on the methods and studies presented in 'Learning Dynamics' and an important point
mentioned in \cite{Marinescu2014} that 'the dynamics implied by multi-agent systems lead to
stochastic behaviour resulting sometimes in undesired effects'. It is clear from the discussion that
this study has important implications for MARL. Specifically, it allows us to consider under which
conditions convergence to an equilibrium is possible. It also allows for directed parameter tuning
to ensure that the resultant game yields desired effects. However, this study is still in its
infancy and there are areas yet unexplored. For instance, it would be important to consider the
challenge presented by Bloembergen et al. in \cite{Bloembergen2015} - stateful games with a
continuous action space. This will by no means be trivial and will require strong assumptions
regarding the learning method used and the structure of the game (two-player, symmetric etc). The
analysis presented in \cite{Letcher2019DifferentiableMechanics} regarding Differential Game
Mechanics would prove as a valuable starting point.

Extending to the large n-player case considers the other end of the spectrum which builds on the
work of Hu in \cite{Hu2019} by extending the dynamical study of the mean-field MARL approach which
is gaining popularity. The models may be extended to the consideration of multiple populations,
asymmetric games and continuous action spaces. Perhaps the results presented in \cite{Bellomo2017}
may prove useful in the consideration of the inter-agent effect on the larger population.

Important questions must be considered in this study, such as the ability of agents to predict the
behaviour of other agents. Chaotic dynamics means that this there are limitations to the degree of
predictability that agents show whilst playing a game. These limitations should be explored so that
algorithms which require predictive behaviour may be benchmarked.

As shown in \cite{Bloembergen2015}, these studies are not only useful as a modality for study but
can also be reverse-engineered to develop learning methods which exhibit desired behaviours. To this
end, I propose considering a control theoretic approach towards multi-agent learning. This would
require extending the work of \cite{Berkenkamp2017} and/or
\cite{Jin2018Stability-certifiedPerspective} to the multi-agent case. Note that
\cite{Jin2018Stability-certifiedPerspective} does actually consider a multi-agent case, but with
independent learners who do not consider the effect that they have on the other agent(s). The
particular extension that I propose is to consider the coupled effect between agents using
replicator dynamics.

\end{document}