### 2020-02-04 Literature Review/Literature Review - Revised/.history/sample_20200203092523.tex
Added Chapter 8

\documentclass[11pt]{report}

\usepackage{matlab-prettifier}
\usepackage[english]{babel}
\usepackage[utf8]{inputenc}
\usepackage{amsmath}
\usepackage{float}
\usepackage{graphicx}
\usepackage{caption}
\usepackage{subcaption}
\usepackage{csquotes}
\usepackage{algpseudocode}
\usepackage{algorithm}
\usepackage[colorinlistoftodos]{todonotes}
\usepackage[margin=0.5in,letterpaper]{geometry}
\def\changemargin#1#2{\list{}{\rightmargin#2\leftmargin#1}\item[]}
\let\endchangemargin=\endlist 
\usepackage{listings}
\def\changemargin#1{\list{}{\leftmargin#1}\item[]}
\let\endchangemargin=\endlist

\usepackage{subfiles}

\title{Literature Review - Revised}

\newcommand\fr[1]{\textcolor{red}{FR: #1}}
\newcommand\TODO[1]{\textcolor{purple}{TODO: #1}}

\author{Aamal Hussain}

\date{\today}

\begin{document}

\maketitle

\tableofcontents

\chapter{Introduction} \subfile{sections/Intro} \label{ch::Intro}

\chapter{Swarms} \subfile{sections/Swarms.tex} \label{ch::Swarms}

\chapter{Dec-POMDPs} \subfile{sections/Dec-POMDPs} \label{ch::Dec-POMDPs}

\chapter{Game Theory} \subfile{sections/GameTheory} \label{ch::GameTheory}

\chapter{Multi Agent Reinforcement Learning} \subfile{sections/MARL} \label{ch::MARL}

\chapter{Control Theory} \subfile{sections/ControlTheory} \label{ch::ControlTheory}

\chapter{Proposed Research} \subfile{sections/Proposals} \label{ch::Proposals}

\bibliographystyle{IEEEtran}

\bibliography{IEEEfull,references}

\end{document}