### 2020-05-20 ChaoticDynamics/WriteUp.tex
Complete Chapter 8 on Swarm Control

\documentclass{article}

\usepackage[margin=2cm]{geometry}
\usepackage{amsmath}

\title{Derivation Write Up}
\author{Aamal Hussain}

\begin{document}

	\maketitle

	\section{Introduction} % (fold)
	\label{sec:introduction}

	\begin{itemize}
			\item For use in Multi-Agent Reinforcement Learning
			\item Requirement to get some desired behaviours out of the system (e.g. conforming to
			predefined values or constraints)
			\item For this, it is required that we are able to predict the outcome of the learning
			behaviour
			\item Work done on determining the dynamics of MARL gives an indication of what
			solutions an algorithm will produce
			\item We can use this, alongside stability theory, to understand the nature of the games
			which result in stable behaviours and which result in chaotic behaviours.
			\item Largely follows the work presented by Sanders et al. However, whilst they consider
			EWA, we consider Q-Learning, which is more popular in RL communities.
	\end{itemize}

	\subsection{Problem Statement} % (fold)
	\label{sub:problem_statement}

	Given a particular learning algorithm and game, is it possible to determine whether the game is
	likely to reach a unique fixed point or exhibit more complex behaviour? What are the factors
	which affect this resulting behaviour?

	% subsection problem_statement (end)

	\subsection{Objectives and Scope} % (fold)
	\label{sub:objectives_and_scope}

	Analyse the stability of Multiagent Q-Learning on iterated games. Brings the work presented by
	Sanders et al to the reinforcement learning community.

	Assumptions:

	\begin{itemize}
		\item Focuses on stateless games
		\item Homogeneous agents
		\item Discrete action spaces, with large number of actions.
		\item Small, finite number of agents.
	\end{itemize}

	% subsection objectives_and_scope (end)

	% section introduction (end)

	\section{Derivation} % (fold)
	\label{sec:derivation}

	We start with the two-agent Q-Learning dynamics as presented by Tuyls et al.

	\begin{subequations}
		\begin{equation}
			\frac{\dot{x}(t)}{x(t)} = \alpha \tau (\sum_{j} a_{ij} y_j - \sum_{i j} x_i a_{ij} y_j)
			+ \alpha \sum_j x_j ln(\frac{x_j}{x_i}) 
		\end{equation}
		\begin{equation}
			\frac{\dot{y}(t)}{y(t)} = \alpha \tau (\sum_{j} b_{ij} x_j - \sum_{i j} y_i b_{ij} x_j)
			+ \alpha \sum_j y_j ln(\frac{y_j}{y_i})
		\end{equation}
	\end{subequations}

	In order to follow the conventions of spin glass theory for the analysis of disordered systems,
	we rescale the system so that the payoff matrix elements are of order $N^{-1/2}$. The motivation
	for doing this is that, along the way, we will take the limit of the number of actions $N$ to
	infinity. However, doing this will result in numerical underflow of the action probabilities 
	(i.e. the probability of each action goes to zero). To compensate for this, we adjust the
	system so that the sum of the probabilities add to $N$. The scaling goes as follows:

	\begin{subequations*}
		\begin{equation*}
			a_{ij} = \sqrt{N} \tilde{a_{ij}}
		\end{equation*}
		\begin{equation*}
			b_{ji} = \sqrt{N} \tilde{b_{ji}}
		\end{equation*}
	\end{subequations*}

	We compensate for this change with
	\begin{subequations*}
		\begin{equation*}
			x_{i} = \tilde{x_{i}}/N
		\end{equation*}
		\begin{equation*}
			y_{i} = \tilde{y_i}/N
		\end{equation*}
	\end{subequations*}

	This gives the original equations as

	\begin{subequations}
		\begin{equation}
			\frac{\dot{\tilde{x_i}}(t)}{\tilde{x_i}(t)} = \alpha \tilde{\tau} \sum_{j} \tilde{a_
			{ij}} 
			\tilde{y_j} -
			\tilde{\alpha} \tau \frac{1}{\sqrt{N}} \sum_{i j} \tilde{x_i} \tilde{a_{ij}} \tilde{y_j}
			+ \tilde{\alpha} \sum_j \tilde{x_j} ln(\frac{\tilde{x_j}}{\tilde{x_i}}) 
		\end{equation}
		\begin{equation}
			\frac{\dot{\tilde{y_i}}(t)}{\tilde{y_i}(t)} = \alpha \tilde{\tau} \sum_{j} \tilde{b_
			{ij}} 
			\tilde{x_j} -
			\tilde{\alpha} \tau \frac{1}{\sqrt{N}} \sum_{i j} \tilde{y_i} \tilde{b_{ij}} \tilde{x_j}
			+ \tilde{\alpha} \sum_j \tilde{y_j} ln(\frac{\tilde{y_j}}{\tilde{y_i}}) 
		\end{equation}
	\end{subequations}

	% section derivation (end)

\end{document}