### 2020-06-03 Documents/ESA/sections/LitReview.tex

\documentclass[.../main.tex]{subfiles}

\begin{document}

    \section{Multi Agent Systems} \label{sec::Multi_Agent_Systems}

    \section{Game Theory} \label{sec::Game_Theory}
    \subsection{Stochatic Games} \label{sec::Stochastic_Games}
    \subsection{Game Theoretic Control} \label{Game Theoretic Control}

    \section{Multi Agent Reinforcement Learning} \label{sec::MARL}
    \subsection{Learning in Two Player Games} \label{sec::Two_Player_Games}
    \subsection{Learning in Stochastic Games} \label{sec::Learning_Stochastic_Games}
    \subsection{Multi Agent Learning Dynamics} \label{sec::MARL_Dynamics}

    \section{Control Theory} \label{sec::Control_Theory}
    \subsection{Distributed Model Preodictive Control} \label{sec::Distributed_MPC}
    \subsection{Approaches to Swarm Control} \label{sec::Swarm Control}

    \section{Swarms} \label{sec::Swarms}
    \subsection{Decision Making in Swarms} \label{sec::Decisions_in_Swarms}

\end{document}