### 2020-06-03 Documents/ESA/sections/ProposedResearch.tex

\documentclass[.../main.tex]{subfiles}

\begin{document}
    \section{Chaotic Dynamics in MARL} \label{sec::Chaos_in_MARL}
    \section{Large Agent Dynamics} \label{sec::Large_Agent_Dynamics}
    \section{Swarm Control through Fields} \label{sec::Swarm_Field_Control}
    \section{Incorporation of Intelligence in Control} \label{sec::Intelligence_in_control}
\end{document}