### 2020-06-03 Documents/ESA/sections/Intro.tex
Added Results to Chaotic Dynamics

\documentclass[.../main.tex]{subfiles}

\begin{document}

    Swarm robotics bases its success upon the collaboration of many agents who contribute towards
    the overall success of the system \cite{HammanFormalApproach}. The union of populations of
    agents who share the same goal can give rise to behaviours that are far more complex than
    individual robots can achieve. The same, of course, is true for humans; whilst an individual may
    be capable of building a house, it takes many more to build a city.

    The ability to exhibit complex

    \section{Problem Statement} \label{sec::Problem_Statement}

    \section{Objectives and Scope} \label{sec::Objectives_and_Scope}

    \section{Methodology} \label{sec::Methodology}

\end{document}