### 2020-06-04 Documents/ESA/sections/ProposedResearch.tex

\documentclass[.../main.tex]{subfiles}

\begin{document}

	\textbf{Major restructuring and rephrasing is required here}

    \section{Chaotic Dynamics in MARL} \label{sec::Chaos_in_MARL}
    The intention of this area of study is to consider the ideas presented in works such as 
    \cite{Sanders2018} and \cite{Galla2011} which consider complex behaviour, including cycles and
    chaos, in certain games using simple reinforcement learning algorithms. These provide a great
    deal of insight into whether or not the game will converge and, if so, to what equilibrium.
    Galla, for instance, shows that adjusting a memory parameter when learning to solve an Iterated
    Prisoner's Dilemma game can shift the equilibrium from one showing purely defective behaviour to
    one showing cycles of cooperations and defection. Sanders et al. generalise this result by
    showing that, for a particular learning algorithm known as Experience Weighted Attraction,
    learning dynamics varies dependent on the mutual effect of two parameters.

    This determination of where RL algorithms will show complex dynamics is important, not only for
    a priori understanding of resultant behaviour, but is particularly important for algorithms in
    which agents aim to predict the behaviour of other agents. For this to be feasible, the learning
    dynamics must not exhibit chaos, otherwise it would be impossible for an agent to make any
    reasonable predictions about the future behaviour of its opponents.

    To this end, the potential avenue for research in this regard is to extend the ideas of Sanders
    et al. and Galla et al. towards games learnt using classical reinforcement learning algorithms 
    (such as Q-Learning with Boltzmann exploration, which seems to be a common choice). The work may
    then be able to extend beyond this, and towards continuous spaces, stateful games and large
    agent populations. This would allow for an understanding of the important parameters which
    define the behaviour of reinforcement learning algorithms and allow for a characterisation of
    the expected resultant behaviour when applying a particular algorithm to a particular game. It
    would also allow for a characterisation of the conditions under which MARL algorithms may
    feasibly be applied, thereby supporting the ability of researchers and engineers to choose their
    payoff matrices and parameters accordingly.
    \section{Large Agent Dynamics} \label{sec::Large_Agent_Dynamics}
    This suggestion aims to build upon the work presented by Leung et
    al. \cite{Hu2019} in which the point is made that the vast
    majority of the work done in this area considers games with a
    finite (usually two) number of agents. This limits the capability
    of the analysis to generalise to much larger agent populations,
    such as swarms. To allieviate this concern, the authors present an
    anaysis of the learning dynamics for a large agent population
    (which they approximate as containing infinite agents) where each
    agent is an independent Q-Learning using Boltzmann
    exploration. The emergent dynamics is given by a Fokker-Planck
    equation which is numerical shown to be a strong approximation of
    the true learning dynamics. However, as the authors point out,
    this is the first attempt at considering such a problem, and
    relies on heavy assumptions.  Possible extensions therefore
    require a consideration of:

    \begin{itemize}
        \item Asymmetric games, in which agents have different
          payoffs, introducing heterogeneity and competition into the
          system. \fr{this is not really a problem  in relation with swarms, is it?}
        \item Stateful Games. \fr{what is the interest of stateful games?}
        \item Heterogeneous populations, in which there are p different populations of agents.
    \end{itemize}

    To the best of my knowledge, in the particular case of $N >> 1$ agent populations, none of the
    above have yet been attempted.

    \section{Swarm Control through Fields} \label{sec::Swarm_Field_Control}
    This study extends the work of Zhang \cite{Zhang2018}, Elamvazuthi \cite{Elamvazhuthi2019} and
	Li et al. \cite{Li2017}. Here, the authors consider the dynamics of swarms through 'drifted
	brownian motion'. In this case, the control input is some scalar field which is generated from the
	environment. Each agent measures this field at their location and uses it to adjust their
	velocity, which eventually results in the entire swarm achieving a desired distribution. This technique
	has been shown, both theoretically and experimentally, to drive swarm systems in a stable
	manner.

	The hypothesis of this section is that the distribution of active particles may
	similarly be controlled through the influence of a scalar field. The dynamics derived by Bellomo
	et al. \cite{Bellomo2017} includes a 'flocking field' which agents can measure. By controlling
	the flocking field, we may be able to drive the swarm to specific
	distributions.

	The contribution that this study presents over those using brownian motion is the inclusion of an
	interaction operator. This governs how agents interact with one another. By accounting for this, we
	are able to control swarms where agents, for instance, avoid collisions with one another. The
	complexity of this interaction term will need to be gradually increased over time. To begin with,
	we may entirely neglect this term. It is here that we will compare the model proposed by Zhang with
	that proposed by Bellomo et al. We may then consider only local interactions, which is typical for
	most swarming systems in the current literature, before then considering non-local interactions.
	This provides the scope for swarms to interact with a greater number of agents. At this
	stage, we may expand our study to consider the presence of interaction domains. This places
	constraints on how agents may interact with one another, allowing for a greater generalisation of
	inter-agent interactions. Ultimately, we would like to influence the interaction term, though this
	is discussed in Section \ref{sec:learning_strategies_through_iterated_interaction}. The questions
	we will consider here are:

	\begin{itemize}
		\item The controllability of the system: Under what conditions is it possible to drive the
		swarm from one configuration to another in a stable manner?
		\item The guarantees of the system: Is it possible to ensure that the agents will not enter
		certain areas of state space, or exceed certain control constraints (e.g. max imum velocity)?
	\end{itemize}

	Some applications of this study are:

	\begin{itemize}
		\item Search and Rescue, in which agents will be required to flock to specific areas (e.g.
		areas with survivors detected).
		\item Construction, in which agents will be required to take on desired formations based on
		the structure to be designed.
		\item Medical Robotics, in which agents are required to target specific areas of the body.
	\end{itemize}

	In conclusion, the results from this study already significantly expand the current capabilities of
	swarming agents by developing control methodologies which can drive the swarm towards desired
	distributions in a stable manner. This is accomplished through controlling the 'flocking field'
	which is indirectly used to adjust the velocity of the agents.

    \section{Incorporation of Intelligence in Control} \label{sec::Intelligence_in_control}

    This section of the study is perhaps the strongest extension to swarm control proposed in this
	chapter, and will likely be the most challenging. Here, we attempt to control the interaction term
	in Bellomo et al's model and we consider that the behaviour of the entire swarm may be influenced
	by decision making on an individual level. This will involve agents learning their strategies
	through repeated interaction. The goal, then, is to design the agent parameters, reward structure
	and control strategy to achieve desired results. To accomplish this will require a unification of
	multi agent learning and control theory. Whilst, at first glance, such an endeavour seems
	incredibly challenging, our work in the predictability of mean field learning (Chapter 
	\ref{ch::DynMARL}) places us in a unique position to tackle the problem. This framework will
	first establish the desired strategy to be taken by the agents and choose the game's reward
	structure so that learning will converge to a stable fixed point. Then by accounting for this
	strategy within the interaction term, we apply the appropriate controls in the flocking field to
	drive the swarm in a stable manner. We will again consider the aforementioned controllability
	and guarantees questions and establish how the incorporation of learning affects these results.

\end{document}