### 2020-06-23 Documents/ESA/main.tex
Updated ESA

\documentclass[11pt]{report}
\usepackage[toc,page]{appendix}
\usepackage{amsmath,amsfonts}
\usepackage{mathtools}
\usepackage{graphicx}
\usepackage{caption}
\usepackage{subcaption}
\usepackage{color}
\usepackage{subfiles}
\usepackage[margin=.8in,letterpaper]{geometry}
\usepackage{fancyhdr}
\usepackage{tikz}
\usetikzlibrary{calc}

\pagestyle{fancy}
\fancyhf{}
\rhead{\thepage}
\lhead{\leftmark}

\begin{document}

\clearpage\thispagestyle{empty}
\begin{titlepage}
    \begin{tikzpicture}[overlay, remember picture]
        \node[anchor=north west, %anchor is upper left corner of the graphic
        xshift=0.5cm, %shifting around
        yshift=-0.5cm] 
        at (current page.north west) %left upper corner of the page
        {\includegraphics[width= 0.5 \textwidth]{Figures/ImperialLogo.jpg}}; 
    \end{tikzpicture}

	\flushright
    \hfill
    \vfill
	{\Huge Early Stage Report \par} 
	\rule[5pt]{\textwidth}{.4pt} \par
	{\Large Aamal Abbas Hussain \\ CID: 01060230}
	\vfill 
    {\Large Department of Computing\\Imperial College London\\Supervisor: Francesco Belardinelli\\
	June 2020 \\}
\end{titlepage}

\subfile{sections/Abstract.tex}

\tableofcontents

\chapter{Introduction} \subfile{sections/Intro.tex} \label{ch::Intro}

\chapter{Literature Review} \subfile{sections/LitReview.tex} \label{ch::Lit_Review}

\chapter{Research Summary} \subfile{sections/ShortTermGoals.tex} \label{ch::ShortTermGoals}

\chapter{Proposed Research} \subfile{sections/ProposedResearch.tex} \label{ch::ProposedResearch}

\chapter{Research Timeline} \subfile{sections/ResearchTimeline.tex} \label{ch::ResearchTimeline}

\bibliographystyle{IEEEtran}

\bibliography{IEEEfull,references}

\begin{appendices}
      \chapter{Research Summary} \subfile{sections/ResearchSummary} \label{app::ResearchSummary}
\end{appendices}

\end{document}