### 2020-06-26 Documents/ESA/sections/LitReview.tex

\documentclass[.../main.tex]{subfiles}

\begin{document}

\textbf{TODO: Mention the parts of literature which we are not considering and explain why: Recent
MARL techniques (including MDRL), Marco Dorigo's work on Swarms. Add in the Learning in Stochastic
Games literature from Schwartz et al. Explain the concepts in Control Theory and how MPC works. Add
in suggestions for applications and papers supporting said applications.}

In this chapter, we explore the current state of the art in
Multi-Agent Systems (MAS). We will consider the main techniques that
have been proposed towards task allocation and control. In this review we cover

\begin{enumerate}
    \item Game Theory: The most well studied model for solving multi-agent decision making problems.
    \item Multi-Agent Reinforcement Learning (MARL): A contemporary technique which requires that
    agents learn and adopt their own strategies.
    \item Control Theory: Methods focusing on low-level manipulation of a system (e.g. the position
    and velocity of an agent) which aims to provide guarantees on its behaviour.
    \item Swarms: Systems whose success are founded on the emergent phenomena exhibited by large
    populations of interacting entities.
\end{enumerate}

There are, of course, a number of techniques which we do not discuss here, as they are not relevant
to the proposals in this report. Most notable is that of Decentralised Partially Observable Markov
Decision Processes (Dec-POMDPs) which is a multi-agent extension of the MDP framework in which
agents aim to maximise a common reward (or utility) function. This method does not scale to large
populations and so is not considered relevant to our study. The interested reader should consult
\cite{OliehoekDecentralizedPOMDPs} for an excellent introduction to Dec-POMDPs.

\section{Multi Agent Systems} \label{sec::Multi_Agent_Systems}

We will keep the problem statement defined in Section \ref{sec::Objectives_and_Scope} in mind as
we define several metrics which a Multi Agent System (MAS) may be assessed against.

\begin{enumerate}
    \item {\em Distributed}. It will often be the case that a multi-agent system must spread out
across a large area or operate in regions where communication is limited. As such it is beneficial
that such a system does not rely too heavily on communication with a central body. The individual
agents must, therefore, be able to make their own decisions and act as independent entities. Aside
from the low communication load, this ensures that the system is not tied to a single point of
failure; in a system governed by a centralised decision maker, any faults in the central body
propagates throughout the group. Of course, this comes at the price of computational load - if the
agents are required to act independently, they must possess the resources to do so. The term
distributed should not be confused with decentralised. The latter refers to a regime in which agents
all communicate with one another before making a decision. Whilst this is preferable to a
centralised decision maker in terms of the requirement for one entity to communicate with all
agents, it still places a high cost on the requirement that agents be in constant communication with
one another. In fact it may pose the additional risk that if any agent in the team fails, the whole
pipeline falls apart. Distributivity, on the other hand, requires that agents make independent
decisions so that, though agents may communicate with each other, they are still able to operate
when this is not possible.
    \item {\em Generality}. MAS come in many different flavours; sometimes a small group of
heterogeneous  (agents with different capabilities) agents may be required to perform a delicate
task whilst in others, it may be required that a large population of homogeneous (all of the same
type) agents perform multiple tasks simultaneously. Indeed the same MAS may have to adapt its
strategy whilst in the middle of a task. It would be a hindrance if the agents were required to be
reprogrammed with a completely different control methodology each time the situation changes
slightly. It would be preferable if the method could apply to all different types of scenarios, with
homogenous or heteregeneous agents, with single tasks or multiple, in any environment with only
minor customisation required. Better still if the agents are able to adapt themselves to the
specifics of the scenario.
    \item {\em Low Computational Load}. As mentioned in the previous chapter, a MAS will often be
used in complex situations, many of which will be too dangerous for human intervention. As such, the
system will be required to carry out its tasks online. For this reason, it is vital that the control
method be one which produces low computational load - especially where the time frame for adjusting
agent behaviour is limited. Monetary concerns also play a part in this metric; it would not do to
impose that every member of the team be equipped with high performance computing power as this would
likely fall outside of the budget of any practical usage.
    \item {\em Robustness}. Particularly for physical robots, a MAS should be able to operate in
situations which present sensor noise, communication noise, latency, and environmental disturbances.
This is rather important in unseen environments where full a priori knowledge of the map cannot be
assumed.
    \item {\em Scalable}. This is mentioned last since the ability for a MAS to scale comes from
its distributivity  and low computational load. Scalability is the capacity for any method of
controlling a MAS to successfully operate as the size of the team increases. This, of course, is
related to the complexity of the algorithms involved. However, it should be noted that it is not
always required that a MAS be formed of dozens (or even hundreds) of agents. Depending on the task
at hand (e.g. robotic surgery), it may be beneficial for fewer, more capable agents to perform the
task, rather than a large population. 
\end{enumerate}

We begin by presenting a subset of the vast array of literature regarding MAS and focus on those
which are relevant to the study at hand. We conclude this chapter by providing some remarks about
how these methods hold against the above metrics.

\section{Game Theory} \label{sec::Game_Theory}

Game Theory has a rich history when considering an understanding of multi-agent systems and it is
impossible to examine MAS without an understanding of Game Theory. These begin in economics but have
found a strong application in computation due to the rising need for distributed systems. Game
Theory, therefore, branches across all of the categories in this chapter although its synergy with
swarms requires development) since Dec-POMDP (Decentralised Partial Observable Markov Decision
Processes) and MARL methods have both base their theory upon the foundations of Game Theory.

\subsection{Partially Observed Stochatic Games (POSG)} \label{sec::Stochastic_Games}

POSGs act as a game theoretic corollary to the Dec-POMDP. Here, an optimal solution to a multi-agent
decision making problem is found by
fixing the strategy (a manner of selecting an action \cite{Rizk2018}) of all agents
except one. The optimal strategy of the chosen agent is then computed. This is fixed so that the
next agent may determine its optimal strategy. This process is iterated over all agents in a process
known as Alternating Maximisation \cite{Ray2010}. By choosing strategies which maximise
a common payoff, a globally optimal solution can be found.

Similarly Bayesian Games is a game theoretic approach with incomplete information (thus it is
sometimes referred to as an `incomplete information game’). This allows for a POSG to be
approximated by a sequence of smaller, more tractable games
\cite{Emery-Montemerlo}. However, iterating through the entire team of agents to
find an optimal solution is a lengthy process and does not allow for immediate actions in dynamic
situations. To address this, \cite{Ray2010} propose to place distinct payoffs on each
agent and to consider team formation based on social hierarchy as well as preferred partners. This
allows the complexity of the problem to be broken down into smaller game units and also provides a
clear order in which the games are to be played. This minimises the interference across robots and
allows for more immediate action to be taken by agents who choose their strategy early on, while
leaving more passive tasks for those later in the sequence.

The Bayesian Game formulation therefore provides a strong candidate for rapid task re-allocation and
dynamic decision making. However, as found by Dai et al. \cite{Dai2018} this is
conditional on an understanding by each robot of the strategy of others, which in turn requires
adequate communication between robots. Therefore, the game theoretical formulation will fail where
communication is not possible amongst team members. However, the addition of heuristics, such as
deep learning, may be able to advance an agent’s ability to recognise the strategies taken by other
members of its team without the need for significant explicit communication.

% \textbf{Add in any recent literature in these fields}

\subsection{Game Theoretic Control} \label{Game Theoretic Control}

Game theory can often be applied to problems of control theory
(particularly where there are multiple agents) to develop robust
controllers which guarantee properties of stability and constraint
satisfaction.

This idea is explored in \cite{Marden2018AnnualControl}. Here, a
zero-sum game is considered in which the players are a controller and
an adversarial environment. The design of the controller must be such
that it is able to drive the system to zero error. To illustrate,
consider the problem of designing a controller for a re-entry vehicle,
as in \cite{Breitner1994ReentryGame}, in which vortices seek to
destabilise the agent. This will allow us to build stable agents in a
much more efficient manner since we can simulate the adversarial
environment and hypothetical scenarios the agent may encounter without
actually encountering them. The same notion is explored by Bardi et al
\cite{Bardi1991DifferentialDisturbances}.

Mylvaganam et al, in \cite{Mylvaganam2017AutonomousApproach}, consider
the $N$-robot collision avoidance problem, similarly from the point of
view of differential game theory. They develop a robust feedback
system for the robots which they show to be able to drive the system
towards predefined targets whilst providing guarantees of interference
from other agents (or lack thereof).  In \cite{MylvaganamASystems},
Mylvaganam also considers a game theoretic control of multi-agent
systems in a distributed manner. Here, agents only consider their own
payoff structure and have limited communication with one another. The
author shows that an approximate equilibrium can be found using
algebraic methods and illustrate the capabilities of the technique
using a collision avoidance example. For the sake of brevity the
numerous contributions that Mylvaganam has made to this field is not
presented here. However, we will conclude with those presented in
\cite{Mylvaganam2014}. Here, the author presents approximate solutions
to a number of differential games, including linear-quadratic
differential games (in which system dynamics are linear functions
whilst payoff functions are quadratic), Stackelberg differential
games, where a hierarchy is induced across the players (a notion was
suggested in the research proposal) and mean-field games, which are
discussed under `MARL'. The importance of the linear-quadratic
differential game is the stability of the solution; solutions for the
Nash equilibria (NE) are admissable iff they are locally exponentially
stable (which the author often shows with the aid of Lyapunov
functions). Approximate solutions to the NE are developed which are
more feasible to calculate online. The author then shows that this is
not simply a theoretical exercise by applying the novel methods
towards multi-agent collision problems and designs dynamic control
laws which guarantee that each agent will reach their desired state
whilst avoiding collision with the other agents. Similarly, the
Stackelberg game is applied to the problem of optimal monitoring by a
multi robot system.

\section{Multi Agent Reinforcement Learning} \label{sec::MARL}

Reinforcement learning extends the Markov Decision Process problem by
considering the case where the reward model is not initially known to
the agent. In a similar manner, Multi Agent Reinforcement Learning
(MARL) extends the Markov Game setting to one where the payoff
structure is not a priori knowledge.

The task of MARL is to determine an optimal joint policy for all
agents across the game. This joint policy may be the concatenation of
all the individual policy or it may just be options for each agent to
take. In either case, optimality is defined through the standard
notions of Nash equilibria and so, in this section we
consider the broad spectrum of methods which attempt to achieve
Nash equilibria.

The largest problem in MARL is the non-stationarity
of the environment \cite{Hernandez-LealA}. In single-agent settings,
it is assumed that the environment is Markovian.  However, this must
be lifted in the Multi Agent setting since other agents in the
environment will be learning concurrently. As such, we must now
consider that the policy for any one agent will depend on the policy
of all other agents.

This section begins with a selection of the foundational methods which
were developed towards solving the MARL problem. The interested reader
may find additional methods and implementations of these techniques in
\cite{SchwartzMulti-agentApproach}.

\subsection{Learning in Two Player Games} \label{sec::Two_Player_Games}

The most fundamental method to learning in Matrix games is the simplex
algorithm. This is a popular method of linear programming (in which
constraints are linear). This will be important in considering more
current methods. A similar consideration is given to the infinitesimal
gradient ascent algorithm, in which the step size converges to
zero. This method guarantees that, in the infite horizon limit, the
payoffs will converge to the Nash equilibrium payoff. Note that this
does not necessarily mean that both agents will converge to a single
Nash equilibrium. This is a particular problem in games where there
are multiple Nash equilibria. However, in practice it is difficult to
choose a convergence rate of the step size and, without an appropriate
choice the strategy may oscillate as shown in the book \cite{SchwartzMulti-agentApproach}. To
address
this, a modified approach is presented by Bowling and Veloso which
incorporates the notion of Win or Learn Fast (WoLF) to produce
WoLF-IGA \cite{Bowling2002}.  WoLF is a notion we will come across
often in MARL and is shown by the authors to converge to always to a
NE. The concern with WoLF methods, however, is that it requires
explicit knowledge of the payoff matrix (which is not so much of a
problem for model based methods) and the opponent's strategy (which is
more of a problem in real-world methods). Finally, the Policy Hill
Climbing method (PHC) is shown to converge to an optimal mixed
strategy if the other agents are stationary (i.e.~are not
learning). However, it is shown that, when this is not the case, the
algorithm again oscillates. The WoLF-PHC adaptation of this method is
shown to converge to a NE strategy for both players with minimal
oscillation.

\subsection{Learning in Stochastic Games} \label{sec::Learning_Stochastic_Games}

Stochastic Games (or Markov Games) form a basis for MARL
settings. However, in this case the agents must learn about the
equilibrium strategies by playing the game, which means they do not
have a priori knowledge of the reward or transition functions. Schwarz \cite{SchwartzMulti-agentApproach}
considers two properties which should be used for evaluating MARL
algorithm: rationality and convergence. The latter simply states that
the method should converge to some equilibrium whereas the former
suggests that the method should learn the best response to stationary
opponents. A similar set of conditions is considered by Conitzer and
Sandholm in \cite{ConitzerAWESOME:}. Schwarz then presents a review of MARL methods (as of
September 2014).

% \textbf{Add in more recent literature r.e. learning in stochastic games and the robustness of
% MARL}

\subsection{Multi Agent Learning Dynamics} \label{sec::MARL_Dynamics}

Multi Agent Learning Dynamics (often referred to as Game Dynamics or
Learning Dynamics) considers the problem of mathematically modelling
Multi Agent Systems who adapt through repeated interact with one
another. This model then serves to be able to predict the evolution of
the system as well as to understand the trajectory of
learning. Typically, this looks at considering whether or not the
method is likely to converge towards a Nash equilibrium. This is
generally a difficult problem to solve
\cite{ShohamMultiagentFoundations} for all but toy problems.
\cite{Letcher2019DifferentiableMechanics} shows that the stable
equilbrium and Nash equilibrium (NE) are not necessarily the same and,
in fact, argue that stable points are more informative than NEs.
Stability provides some guarantees against the stochastic nature of
the environment since a stable equilibrium will always be returned to
even after perturbations. This feature is extremely important in Safe
and Trusted AI as it provides guarantees against undesired behaviour
in real world environments.

The area of dynamics which has shown most promise in Multi-Agent
Reinforcement Learning is that of evolutionary dynamics. This draws
from the principles of Evolutionary Game Theory (EGT) which considers
similar assumptions to that of MARL: agents are no longer required to
be rational and play the game optimise their expected return through
repeated play. Importantly, players have no knowledge of the others'
payoffs \cite{Tuyls2006AnGames}. In \cite{Tuyls2006AnGames}, Tuyls et
al.  determine the relation between the replicator dynamics concept of
EGT (a differential equation defining the evolution of the proportion
of a subgroup in an evolving population) and Q-Learning using
Boltzmann probabilities as Q-values. The result was a dynamics
equation which describes, for each action, the evolution of its
selection probability which could even account for random
exploration. There have since been a number of works which apply the
same insight into different game types and MARL algorithms. In
\cite{Bloembergen2015}, these are broken into the following categories

\begin{itemize}
    \item Stateless games with discrete actions. Here, stateless refers to the idea that the game is
    static and so the environment has no impact on the result. The aforementioned result 
    \cite{Tuyls2006AnGames} fits into this category.
    \item Stateless games with continuous actions. These consider more realistic MARL than the
    previous category by replacing each agent's strategy vector with a probability density function 
    (pdf) over a continuous action space.
    \item Stateful games with discrete actions. This mostly considers stochastic games, where there
    are multiple states with probabilistic (usually Markovian) transitions between them. However,
    extensive form games, which considers more complex phenomena such as sequential moves and
    imperfect information, are also briefly mentioned.
    \item Stateful games with continuous actions. This is one of the more realistic assumptions
    considered. However, the authors point out that this area is yet to see results, leaving it open
    for possible research.
\end{itemize}

Though the above have seen success through experimental validation in games with 2 players and (in
general) 2 to 3 discrete actions, recent work has begun the consideration of improving the models
towards more complex scenarios with larger agent populations. A recent example of this is 
\cite{Hu2019}, in which Hu uses a mean field approximation to model the dynamics of a population of
Q-Learning agents. As a reminder, mean field (MF) approaches in MARL consider that an agent updates
its strategy based on the mean effect of the population. The result is a system of three equations
which describes the evolution of Q-values over a large population in a symmetric bi-matrix game.This
presents an important first step in modelling the learning dynamics of large agent populations and
has the scope to be expanded to systems of asymmetric games, heterogeneous populations and stateful
games.

An advantage of determining the evolutionary dynamics of learning is that it can describe the
expected behaviour in different game settings. This is particularly important to understand the
convergence of the methods; certain games will often show cyclic behaviour even with the existence
of a strict NE. For instance, Imhof et al. show in \cite{Imhof2005} that a repeated prisoner's
dilemma game results in cyclic behaviour when considering the options of cooperation or defection.

\section{Control Theory} \label{sec::Control_Theory}

The control theoretic perspective considers generating a set of control laws for the system. These
are chosen with the aim to satisfy certain properties. The main properties are

\begin{itemize}
    \item {\em Stability}: a system will return to the desired setpoint (or within a neighbourhood) if
    perturbed.

    \item {\em Robustness}: a system will perform its function in the presence of uncertainty and
    noise

    \item {\em Optimality}: the system will achieve some defined function

    \item {\em Feasibility}: the controller will always be able to generate a control law which
    satisfies the desired properties.
\end{itemize}

There are a number of approaches towards control systems. However, the interest of this review lies
in multi-agent systems which must operate in the face of uncertainty. As such, we focus on
stochastic and distributed control. The literature of the control community is vast and can be
divided into a number of sub-fields. We consider Distributed MPC (DMPC) and Swarm Optimal Control as
those which are particularly relevant to this study and review them in this section.

\subsection{Distributed Model Predictive Control} \label{sec::Distributed_MPC}

Rawlings et al. \cite{rawlings2017model} divide the problem of distributed control into four distinct
categories: decentralised control, non-cooperative control, cooperative control, and centralised
control. The last of these is not considered in the text since it only considers the case in which a
centralised controller has access and can manipulate multiple agents at once. This immediately
violates the desired property of distributivity.

Decentralised control is the scheme in which agents do not have information about the actions of
other agents and can only optimise for their own objective. This has the advantage of requiring no
communication, but can often lead to poor performance when the agents are strongly coupled as each
agent’s model is incomplete. In the non-cooperative setting, each agent optimises their own loss
function whilst treating the others’ actions as a known disturbance. In this setting, each agent has
knowledge of the others’ control laws and their effect. In this case, the agents must communicate
their intended actions to one another and iterate to achieve a consensus (or Nash equilibrium). This
is the same for cooperative control, except that the loss function is now shared across the team.

Of all of these systems, cooperative control has found greatest applicability in autonomous systems
\cite{Negenborn2014}, perhaps due to the favourable stability properties \cite{Venkat2006}. One
particularly strong application of this system is in vehicle platooning. Here, self-driving cars or
unmanned aerial vehicles (UAVs) must move in a certain formation without colliding into one another.
A number of examples can be found such as in \cite{Liu2019, VanParys2017, Zheng2017}. Distributed
MPC provides the advantage that guarantees can be placed regarding coupled constraint satisfaction
and feasibility.

The advances in stochastic and robust MPC, however, do lend themselves towards revisiting the
capabilities of decentralised control. Recall that, in this scheme, the agents cannot communicate
with one another. However, advances in MARL show us that this disadvantage can be made less
prevalent if each agent has a model of the other \cite{Foerster}. To this end, a methodology such as
presented in \cite{Heirung2019} may prove beneficial in the decentralised scheme. Here, the system
chooses amongst a family of system models when choosing its control laws. Similarly, agents who
determine (or perhaps learn) models of the other agents may be able to leverage this information,
minimising the model error when optimising.

% \textbf{Include recent updates in here}

\section{Swarms} \label{sec::Swarms}

Swarm systems comprise of a population of (typically) homogeneous
agents who are able to organise themselves into formations using a
series of simple local interactions with their neighbours
\cite{Couceiro2015}. Whilst the individual agents are generally
simplistic, the collective behaviour may exhibit complex phenomena
emulating systems observed in biological organisations such as bee or
ant colonies \cite{Sethi2017}. Hybrid algorithms such as in
\cite{Gao2018} show an accelerated performance in reaching globally
optimal solutions in search-based tasks. The advantage of many swarm
algorithms is that they are based on local interactions and so are
incredibly scalable \cite{Rizk2018}.

\subsection{Approaches to Swarm Control} \label{sec::Swarm Control}

Recent work has seen the advancement of swarms controlled in stochastic environments. This is
particularly important for swarm intelligence in robots; many systems are developed with the
motivation of search-and-rescue, in which the robot swarms will have to operate in environments
where accounting for uncertainty is critical. To this end, swarm systems have seen the advent of
stochastic control. In this, the swarm is modelled as a diffusive system using a stochastic
equation, most often the Kolmogorov Forward Equation \cite{Hamann2008}. In \cite{Hamann2008}, the
author shows that this equation can be derived by considering local trajectories of smaller subsets
of the swarm. Elamvazhuthi and Berman \cite{Elamvazhuthi2019} extend this idea to other stochastic
models for swarms, importantly considering an advection-diffusion-reaction model which allows for
hybrid agents to switch between different modes of operation.

The above models come under the term ‘mean-field models’ \cite{Elamvazhuthi2019b}, a series of
equations for stochastic forward processes (such as swarm foraging) in which, as the number of
agents tends to infinity, the true macroscopic motion of the system tends to these equations.
Importantly, however, these stochastic equations allow for an analysis of the swarm, as well as the
ability to develop control laws. In \cite{Li2017}, Li et al. show that these models can be used to
develop control laws for robots in a swarm and drive them towards a target distribution. The method
is shown to perform accurately both in simulation and on real robots. The advantage is that
guarantees can be placed on the convergence and stabilisabiilty of the swarm towards the desired
distribution. However, there appears to be significant scope to expand upon this methodology from a
safety perspective. To begin with, the method does not consider inter-agent interactions and
therefore does not formally guard against collisions. In \cite{Inoue2019}, a similar problem is
considered, although collisions are avoided by having the robot simply move in the opposite
direction when encountering an obstacle. It is here that the mean-field models are not as strong
since they do not take these local interactions into account. It must be noted, however, that the
authors of \cite{Inoue2019} have identified this problem and are currently working towards its
incorporation into the model.

It would, therefore, be of particular interest from a safety perspective to consider agent
interactions. A starting point may be \cite{Bellomo2017} which extends the dynamical model of a
swarm system to include agent interaction. This presents a first step towards considering the swarm
as composed of intelligent agents rather than mindless particles and, as the authors suggest,
presents the possibility of applying game theoretic approaches towards swarms, which also gives the
ability to consider heterogeneous swarm systems who interact with one another through repeated play.
Similarly, it would allow for a stronger control perspective on swarm systems such as presented in
\cite{Borzi2015}, in which a model predictive control (MPC) scheme is presented for a
leader-follower swarm system to achieve given tasks.

\subsection{Decision Making in Swarms} \label{sec::Decisions_in_Swarms}

% \textbf{Include Marco Dorigo and Recent AAMAS literature in here}

The reduction of the complexity in interactions between agents also
allows for the robots to perform other calculations on board. In
\cite{Pini2011TaskSelection}, Pini et al.~leverage this by considering
adaptive task partitioning across swarms. This allows a swarm, in a
decentralised manner, to deliberate whether to partition a task into
its sub-tasks or to perform the task in its whole. As of now (to the
best of my knowledge) the problem of partitioning general tasks into
its $N$ sub-tasks is unexplored. This, however, highlights another
advantage of swarm systems; they are readily divided into sub-groups
(as in \cite{Zahadat2016DivisionInhibition}) to perform a
{\em divide-et-impera}
%(divide and conquer)
approach to solving problems
\cite{Pini2011TaskSelection}).

Furthermore, swarm systems may be designed in a leaderless manner and
so do not require the use of a central controller
\cite{Couceiro2015}. This presents the advantage that the system can
rapidly adapt to the loss of agents or separation of groups throughout
the task. However, the assumptions made regarding the homogeneity of
individual agents and the simplicity of their local interactions
result in significant limitations placed on the complexity of the
tasks that swarm systems can accomplish.

\section{Discussion} \label{sec:remarks}

Each of the above methods present different advantages in their
approach towards MAS, as well as a particular set of disadvantages.

Game Theoretic methods have long provided a rigorous approach for
modelling the behaviour of any MAS. It places no limitation on the
heterogeneity of the agents or the task demanded. An appropriate
choice of payoff matrices will account for both of these
specifications. Recent advances in POSGs have lifted the strong
assumptions on prior knowledge that was required to achieve a Nash
Equilibrium and Game Theoretic Control allows for such methods to be
applied in continuous state spaces. Whilst the latter shows particular
promise, game theoretic methods still fall short in terms of
distributivity as they are heavily reliant on communication with a
centralised decision maker to supply rewards, or within the team to
determine each others' actions. They also require a priori knowledge
of the situation in order to set up payoff matrices and so cannot be
regarded as robust to environmental uncertainty.

Multi-Agent Reinforcement Learning builds on the latter two
shortcomings of game theory by allowing the environment to supply
rewards to agents who adapt their behaviour accordingly. MARL is
proving to be a powerful method to control MAS, and is rapidly
improving, but cannot yet be regarded as robust. Studies repeatedly
show that small alterations or noise in the environment can result in
large deviations in behaviour. Furthermore, few guarantees can be
placed on the resulting behaviour of a MARL system, though progress is
being made in closing this gap.

Control Theory, on the other hand, takes pride in the guarantees
placed on the MAS such as: controllability, stability, and
robustness. However, control techniques are often centralised or
decentralised. Distributed methods of control is still an evolving
field, particularly in terms of robustness to uncertainty. In order to
establish guarantees, assumptions are also made on the specifics of
the MAS, and so control theoretic methods have yet to evolve into
general, heterogeneous systems with any given task.

Swarm systems have their strongest advantage in scalability and, in
fact, swarm techniques are built to scale due to the distributivity
and low computational load placed on each agent. This has often come
with the assumption that agents in the swarm act as simple,
homogeneous, particles who follow very simple rules of
interaction. Recent work is showing that this is not a necessary
assumption and, in fact, swarms can leverage heterogeneity and agent
capabilities to ensure robustness and generalisation.

It is clear that each method presents its own set of advantages and
disadvantages. In fact, the greater contributions are made from a
unification of techniques (e.g., game theoretic control) which leverage
the strengths of one technique to mitigate the drawbacks of
another. To this end, we consider the unification of swarm control and
multi-agent learning. The aim is to unify the strengths of swarm
theory, MPC and game theory to study populations of agents who can
make decisions on an individual basis but perform tasks through large
scale collaboration. In the subsequent chapters, we propose a line of
research which explores the properties of a system as well as
establishing practical control methodologies along the way.

\end{document}