### 2020-06-29 Documents/ESA/sections/Abstract.tex
Included Abstract

\documentclass[.../main.tex]{subfiles}

\begin{abstract}

	This project aims to study the evolution of a swarm of intelligent `active particles' under the
	influence of Model Predictive Controls. We specifically aim to investigate the learning dynamics
	of a population of agents following a Q-Learning approach. We also aim to study the evolution of
	such a swarm under Model Predictive Controls generated by field particles. The goal of this
	study is to bridge the gap between swarm theory and multi-agent reinforcement learning to enable
	the study of complex populations of agents, who are capable of independent decision making, as
	well as producing control methodologies to enable the safe operation of intelligent swarms.

\end{abstract}