### 2020-11-09 Documents/ChaoticDynamics/Characterisation of Dynamics/main.tex
Added Research on Network Games

\documentclass{article}
\usepackage[margin = 2cm]{geometry}

\title{Hello World}
\author{Aamal Hussain}
\date{}

\begin{document}
    \maketitle

\end{document}