### 2020-11-09 Documents/Network Games/ResearchCompilation.tex

\documentclass{article}
\usepackage[utf8]{inputenc}
\usepackage{geometry}
\geometry{margin=2cm}
\usepackage{xcolor}

\newcommand{\ah}[1]{\textcolor{blue}{#1}}

\title{Controlled Swarms through Learning on Network Games: Research Compilation}
\author{Aamal Hussain}
\date{}

\begin{document}
    \maketitle

The goal of this revision is to re-evaluate the assumptions and guarantees placed on the
following problem: to what extent is it possible to control the behaviour of a large group of agents
subject to partial information. In particular, we seek to provide theoretical guarantees concerning
the behaviour of these agents from the point of view of safety. It is, therefore, required that we
ensure that the tools we are using are suited to such an analysis. In this spirit, this document
outlines an overview of the research into the game-theoretic and learning approaches towards
learning for large systems of agents.

\subsection*{Network-aggregative games are a stronger model for swarm systems than mean-field}

In particular, I focus this research on the problem of network-aggregative games \cite{Parise}. In this
class of game, we consider an aggregative (large population) system which has an underlying
communication network, i.e. agents adapt their behaviour dependent on the behaviours of their
neighbours in the network, but their knowledge of the other agents is limited (or even
non-existent), This form of game appears to be better suited to the problem of swarms than the
mean-field model, in which an agent would be required to have knowledge of the average behaviour of
the entire population \cite{Paccagnan2019}. This assumption is infeasible for a fully distributed
system, unless the agent is somehow able to communicate with the entire population and perform the
necessary calculations to find the average behaviour of the population.

In addition, the research on learning on network games (sometimes referred to as graphical games
\cite{Li2017}) is a topic which is increasingly being turned to from the
learning and game-theoretic communities alike. The major classes of learning algorithms which have
been explored in this context are: Fictitious Play \cite{Ewerhart2019}, Q-Learning 
\cite{Vamvoudakis2017} and No-Regret \cite{Nagarajan2018}. All
of these continue to undergo rigorous theoretical analysis from the game theoretic and computational
learning communities.

In this spirit, important facts about learning on networks have been established for each of the
learning algorithms. This includes, most importantly, convergence to Nash Equilibria. For each of
these classes of games, the conditions required for this behaviour are established in: 
\cite{Ewerhart2019} for Fictitious Play, \cite{Nagarajan2018} for no-regret, and
\cite{Li2017} for Reinforcement Learning. For the last of these, however, it is important to note that,
whilst the paper presents an algorithm for finding the NE, the focus of the study concerns forming
control policies based on the NE. This is not a disadvantage of the paper but rather an advantage:
namely that the learning community tends to be more invested in the application of the algorithms
towards real-world problems than the game-theoretic community.

The existence of these results means that there is a basis towards forming safe learning systems on
network games, since we can study the convergence towards an equilibrium. In addition, this class of
game can, particularly for Fictitious Play and Q-Learning, be extended towards network-aggregative
games, thereby allowing its study for larger populations. Such a study can be found in \cite{Parise},
in which the authors produce two NE seeking algorithms which, although they do not call it so,
is remarkably similar to Fictitious Play. This provides a useful (and perhaps only) study on
network-aggregative games which can be used to compare future work against.

\subsection*{The work on learning on network games is quite recent.}

To qualify the above statements, it is important to note that the work in learning on network-games
and, in particular, network-aggregative games is quite new. Indeed the publications listed in this
review, which represent some of the major breakthroughs in network games, all fall within the span
of 2015-2020. Whilst these, of course, build on previous work, it is important to note that the
specific problem of learning and control on network-aggregative games is an open problem. This is
because network games can be considered to be one of the more generalised forms of game since it
imposes an underlying communication network across the system. Indeed, a game in which an agent acts
according to all of the other agents in the group can simply be considered a fully-connected
network. There are, however, still further generalisations that could be made. For instance,
learning on  weighted networks has received some attention \cite{Bayer2019} but still
requires maturity. Furthermore, directed, or asymmetric networks and time-varying networks are
topics left, as of yet, unexplored. This exposition is not to suggest directions for our own
research, but merely to present the fact that the problem of learning on network games is a valuable
one to consider due to its generality.

\subsection*{Then comes the question of which learning algorithm to study.} % (fold)

Hopefully, I have made a strong argument for why network-aggregative games are a good class of games
to focus the study on for the problem of learning swarm behaviour. It then becomes incumbent on us
to choose a class of algorithm to consider. Whilst each algorithm presents its own set of pros and
cons, I argue that choosing the algorithm is merely a matter of considering the assumptions that
would be realistic to place on our system. As you mentioned in our discussion, if it is the case
that the problem we consider satisfies the assumptions for one of these classes, it becomes
necessary that we choose this class to study.

I would argue that the major conditions that we need to place on our system is that the agents have
no knowledge of the behaviour of the other agents and must, instead, infer that behaviour. Swarm
systems are built on the assumption that an agent is able to interact with, and observe, only a few
agents in the population. This would typically include its nearest neighbours and, if the situation
demands, a leader or set of leaders.

The other typical assumption placed on learning systems is that the agent does not have access to
the utlity function, from which it derives its behaviours. Such a problem would occur in games which
need to be repeated over time (e.g. traffic routing problem) or for the design of a system which is
aimed to be generalisable. However, it is unlikely that, for the problem of controlling a swarm,
this assumption would be required in general. For the most part, a control problem is designed for a
specific task in mind (e.g. coverage or foraging) and so the objective functional is fixed.
Furthermore, as long as complex gradients or Hessian inverses do not need to be calculated, it can
be assumed that the agent is capable of carrying out the necessary calculations in order to perform
its task. Otherwise, it would be required that a centralised controller, who has access to the cost
function and can carry out the optimisation, would need to communicate with each agent to identify
its behaviour and deliver its next action. Whilst there is nothing inherently wrong with this
method, such a system can no longer be called `distributed'.

With these in mind, then, I suggest that the Fictitious Play (FP) and (perhaps less so)
Reinforcement Learning (RL) algorithms would be appropriate classes of games to study. This is
because no-regret learning considers that the agent is aware of the actions of his/her neighbours
and therefore, does not fulful our requirements. I argue that FP is a stronger class for this
problem description since it is assumed that the agent knows her objective function and explicitly
tries to learn the behaviour of her neighbours. On the other hand, RL does not require knowledge of
neighbours or of the objective function but implicitly attempts to learn both of these by evaluating
the actions. Whilst Q-Learning is more general, then, it lacks the explicit modelling aspect that FP
offers and, as such, its analytical basis is not as strong. I propose, therefore, that a study of FP
on network-aggregative games provides the strongest model of swarm systems. I am not, however,
ruling out the option of looking at both of these algorithms and comparing their performance and
guarantees.

\subsection*{Conclusions}

In summary, I propose that, for the problem that we are considering, namely the problem of
controlling swarms of intelligent active particles, the strongest model to study would be that of
Fictitious Play (and/or RL) on network-aggregative games. This suggestion is due partly on the
similarity between the assumptions placed on this class of games and that of swarm systems as well
as the growing attention that this class of games has received from an analytic perspective. I
believe that the best way to proceed would be to extend the work of \cite{Ewerhart2019} with the
assumptions made in \cite{Parise}. A point which I would like to discuss with you is whether it would
be to attempt a similar extension of the work in \cite{Li2017} or \cite{Vamvoudakis2017}
(i.e. that of RL) and provide a comparison of the results. Regardless, both of these would lend
themselves well to considering the problem of distributed control as shown in 
\cite{Vamvoudakis2017} and \cite{Bianchi, DePersis2020}. Both of these allow for us to
consider the physical side of the problem by imposing that the learning algorithm must respect the
physical dynamics of the agents involved.

I believe this to be a stronger suggestion towards the problem of Active Particle Control than that
which was suggested in the ESA as it allows for us to directly build on our current work with only a
minor modification directing our assumptions closer to the ones required by our problem. It also
allows us to bypass the section of controlling a swarm through PDEs and instead narrows our focus on
game-theoretic learning methods.

\bibliographystyle{IEEEtran}

\bibliography{IEEEfull,references}

\end{document}