### 2020-11-25 Documents/ChaoticDynamics/aamas2021-formatting-instructions/Responses to Reviews for AAMAS.txt
edits on the rebuttal

Review 1:

We thank the reviewer for their comments. For the questions:

1. We aim to focus our study on all normal form games. As you point out, whilst the strategy space of the agents includes a finite set of actions, they are able to play mixed strategies as well as pure ones. The goal of the paper is to identify the evolution of the strategy as the agents learn. Thank you for pointing this out, we will clarify it in revision.

2 and 3. These are excellent points and we are, in fact, using the results of this study to consider how to 'steer' the learning towards stable equilibria. As you mention, there are cases in which there are multiple stable equilibria. In such cases, we have not yet determined a closed form method for steering the system towards a given one. This is an important avenue for future work.

Finally, on the links to evolutionary dynamics (weaknesses), we notice that by deriving the continuous time dynamics of Q-Learning, [20] arrives at a system similar to the replicator dynamic. Our goal in this paper is solely focused on understanding the stability of these dynamics.

Review 2:

Thanks for the insightful remarks. Indeed, many of the techniques we apply are inspired by statistical physics. We believe that these methods can be fruitfully applied to learning on games (as discussed in the Related Work), as the latter makes use of many random variable: the strategies of the agents, the payoffs, etc. We aimed, as much as possible, to focus the paper on the main, theoretical contributions, while providing details about the derivation in the supplementary material. We thank you for taking the time to look at the supplementary material with such rigour.

As you mention, one of the main contributions of the paper is equation (13), which establishes the region of stability. To validate this equation, we tested each of its predictions, including the rather remarkable ones that independent Q-Learning rarely converges, and payoff correlations make no difference to stability.

As regards the style of presentation, we are aware that normally, papers at AAMAS present results in a theorem-proof format. However, we felt that this format would not be appropriate for the results proved here, as the statement of each result would just amount to the last equation in the derivation.

Review 3:

We thank the reviewer for the several points they raise, they will certainly be useful to improve the paper, as well as to provide some clarification.

The reviewer mentions that there is a disconnect between the theory and the empirical  evaluation. 
Actually, the theoretical result (13) makes a number of predictions including: the decrease of stability for increasing alpha or tau, the (rather surprising) lack of dependence on Gamma (the correlation between payoffs), and, for the limiting case of an infinite action set and a large number of players, that the majority of the phase space is unstable. There are, however, practical restrictions when conducting experiments with a large action/player set. We do however, present a number of empirical results that show the overall trend. Consider, e.g., Fig. 6 for 2-6 players, which shows the trend is towards the theoretical predictions. It is in this sense (FB: in what sense?) that the size of the unstable region is overestimated by the theory, a point which was also found in [15]. The message of the paper is to highlight the dependence of stability to these parameters, and argue that it is required for algorithms to be analysed with stability in mind.

As regards cyclic behaviour in the Prisoner's Dilemma, the study we reference is actually different to that which Figure 1 depicts and, in fact, discusses a different learning algorithm (EWA) to that of the figure (Q-Learning). We will clarify this distinction.

Method:

As discussed at the beginning of Sec. 3.1, i_mu (resp. i_kappa) refers to the action taken by player mu (resp. kappa). This allows us to forego using different letters for different players.

The point regarding scaling the system was given briefly on p. 4 to conserve space. A detailed discussion can be found in Section S2.2 of the supplementary material.

You are correct in saying that payoffs were already in a general form. To then analyse the system we parameterise these payoffs by 'Gamma' which measures their correlation. To study the expected behaviour of a system which would result from a given Gamma, we average over all the possible realisations that this choice could take. This means that we can focus on how, on average, the payoff correlations affect stability rather than look at all of the (infinite) possible variations of payoff matrices.

Experiments:

The value of tau was chosen for two reasons: 
1) To remain consistent with the results in [15], thereby justifying the comparison made in the discussion. 
2) As our results show, instability increases with tau. Hence, for higher values of tau, the game is trivially unstable everywhere.

Your reading of 'variance of *actions*' is correct, it denotes how much the actions vary about the fixed point. We will state explicitly how x_i(t) should be read.

TODO: Make the points in the first paragraph, PD and Game assumptions as brief as possible. Respond to comments on notation, scaling the system and variance of the actions